\large How resource abundance and stochasticity affect animals' spatial needs

### 2022-11-08 writing/manuscript.Rmd
added refs

\noindent The amount of resources an organism is able to access is a strong determinant of its probability to survive and reproduce. Resource limitations can cause individuals to experience a negative energetic balance, which can then result in lower fitness [@hou_cold_2020; @le_bot_fishery_2019], changes to community structure [@haney_tipping_2018; @burson_competition_2018; @riotte-lambert_environmental_2020; @ghislandi_resource_2018], altered physiology [@dai_pra_ground_2022; @rocha_life_2021; @le_bot_fishery_2019; @wessling_seasonal_2018], lower chance of reproduction [@stefanescu_timing_2021; @schmidt_interplay_2020; @le_bot_fishery_2019], or even death [@foley_severe_2008; @berger_climate_2018]. Thus, many organisms will change their behaviors in response to changes in local resource abundance. Some species may respond to fluctuations in resource abundance by, if possible, switching to other food sources [@steinmetz_effects_2021; @le_bot_fishery_2019], reducing energetic costs by reducing body heat [@schmidt_interplay_2020] or entering hibernation or torpor [@fjelldal_nightly_2021; @mohr_cellular_2020; @boyles_optimal_2020], but movement represents one of the most readily available traits that species can adjust, whether this be by modifying their home range [@bista_effect_2022; @lucherini_habitat_1996; @relyea_home_2000], migrating [@middleton_green-wave_2018; @geremia_migrating_2019], dispersing [@refs], or moving nomadically [@teitelbaum_beyond_2019; @nandintsetseg_variability_2019].

While there are many other causes of animal movement, the relationship between animal movement and resource abundance has been of interest to biologists for over half a century. @burt_territoriality_1943 considered the search for food as the primary driver for animal movement within an animal's home range. Three decades after, @southwood_habitat_1977 suggested change in resource abundance is a strong determinant of how animals decide when and where to live and reproduce. Two years later, @harestad_home_1979 proposed that the simplest hypothesis for the relationship between an animal's home range size is inversely proportional to resource abundance. More recently, various researches reported that home range size decreases linearly with resource abundance [desert mule deer: @relyea_home_2000; red panda: @bista_effect_2022]. Others demonstrated that spatial needs decrease nonlinearly as habitat productivity increases, with stronger decreases when resources are scarce [e.g., @herfindal_prey_2005; @nilsen_can_2005]. Additionally, an animal's spatial requirements depend strongly on its metabolic rate and diet, as demonstrated over three decades ago by various scientists, including @lindstedt_home_1986 and @reiss_scaling_1988. Thus, there are many drivers of animal movement and many causes of variety between behaviors, and we currently lack a unifying framework for quantifying the effects of resource abundance and unpredictability on animals' spatial needs.

@@

@@

```{r moving-window-examples, fig.cap="Seven-day home range size of a tapir ($Tapirus~terrestris$) in response to changes in mean and variance in resource abundance. (a.) GPS tracking data of the tapir. (b.) Mean resource abundance esimated as the mean NDVI at the locations visited by the tapir. (c.) Varince in resource abundance esimated as the average variance in NDVI at the locations visited by the tapir. (d.) Estimated home range size during each seven-day period, based on 50\\% (bold) and 95\\% (thin) utilization quantiles. (e.) Effect of resource abundance on home range size. (f.) Effect of resource unpredictability on home range size. The effects in panels e and f were estimated using generalized linear models with Gamma cinditional distributions. The tapir movement data corresponds to the individual named \"Anna\" from the Cerrado sample of Medici $et~al.$ (2022).", out.width='100%'}
knitr::include_graphics('../figures/tapir-example.png')
```

>>>>

\noindent The amount of resources an organism is able to access is a strong determinant of its probability to survive and reproduce. Resource limitations can cause individuals to experience a negative energetic balance, which can then result in lower fitness [@hou_cold_2020; @le_bot_fishery_2019], changes to community structure [@haney_tipping_2018; @burson_competition_2018; @riotte-lambert_environmental_2020; @ghislandi_resource_2018], altered physiology [@dai_pra_ground_2022; @rocha_life_2021; @le_bot_fishery_2019; @wessling_seasonal_2018], lower chance of reproduction [@stefanescu_timing_2021; @schmidt_interplay_2020; @le_bot_fishery_2019; @douglas_relative_2014], or even death [@foley_severe_2008; @berger_climate_2018]. Thus, many organisms will change their behaviors in response to changes in local resource abundance. Some species may respond to fluctuations in resource abundance by, if possible, switching to other food sources [@steinmetz_effects_2021; @le_bot_fishery_2019], reducing energetic costs by reducing body heat [@schmidt_interplay_2020] or entering hibernation or torpor [@fjelldal_nightly_2021; @mohr_cellular_2020; @boyles_optimal_2020], but movement represents one of the most readily available traits that species can adjust, whether this be by modifying their home range [@bista_effect_2022; @lucherini_habitat_1996; @relyea_home_2000; @yu_artificial_2022; @arechavala-lopez_common_2019], migrating [@middleton_green-wave_2018; @geremia_migrating_2019], dispersing [@refs], or moving nomadically [@teitelbaum_beyond_2019; @nandintsetseg_variability_2019].

While there are many other causes of animal movement, the relationship between animal movement and resource abundance has been of interest to biologists for over half a century. @burt_territoriality_1943 considered the search for food as the primary driver for animal movement within an animal's home range. Three decades after, @southwood_habitat_1977 suggested change in resource abundance is a strong determinant of how animals decide when and where to live and reproduce. Two years later, @harestad_home_1979 proposed that the simplest hypothesis for the relationship between an animal's home range size is inversely proportional to resource abundance. More recently, various researches reported that home range size decreases linearly with resource abundance [desert mule deer: @relyea_home_2000; red panda: @bista_effect_2022]. Others demonstrated that spatial needs decrease nonlinearly as habitat productivity increases, with stronger decreases when resources are scarce [e.g., Canadian lynx: @herfindal_prey_2005; carnivores: @nilsen_can_2005; tigers: @simcharoen_female_2014; ferruginous hawks: @watson_ferruginous_2020]. Additionally, an animal's spatial requirements depend strongly on its metabolic rate and diet, as demonstrated over three decades ago by various scientists, including @lindstedt_home_1986 and @reiss_scaling_1988. Thus, there are many drivers of animal movement and many causes of variety between behaviors, and we currently lack a unifying framework for quantifying the effects of resource abundance and unpredictability on animals' spatial needs.

@@

@yu_artificial_2022: rat snakes move to artificial structures for better prey availability and better thermoregulation

Although it is generally acknowledged that animals will modify their behavior in response to human activity [@rogala_human_2011] as well as changing conditions [@walther_ecological_2002; @chevin_adaptation_2010; @sih_evolution_2011], a fine-scale understanding of behavioral responses is lacking [@sih_evolution_2011; @beever_behavioral_2017].

@@

```{r moving-window-examples, fig.cap="Seven-day home range size of a tapir ($Tapirus~terrestris$) in response to changes in mean and variance in resource abundance. (a.) GPS tracking data of the tapir. (b.) Mean resource abundance esimated as the mean NDVI at the locations visited by the tapir. (c.) Varince in resource abundance esimated as the average variance in NDVI at the locations visited by the tapir. (d.) Estimated home range size during each seven-day period, based on 50\\% (bold) and 95\\% (thin) utilization quantiles. (e.) Effect of resource abundance on home range size. (f.) Effect of resource unpredictability on home range size. The effects in panels e and f were estimated using generalized linear models with Gamma conditional distributions. The tapir movement data corresponds to the individual named \"Anna\" from the Cerrado sample of Medici $et~al.$ (2022).", out.width='100%'}
knitr::include_graphics('../figures/tapir-example.png')
```
